---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  echo = TRUE
)
```

# opennemr

<!-- badges: start -->
<!-- badges: end -->

## Overview

The goal of {opennemr} is to provide simple access to the OpenNEM API via R.

The OpenNEM project (https://opennem.org.au) aims to make Australian energy 
network data more accessible to a wider audience through a website portal and 
data access API’s and tools.

Through the webport and API, OpenNEM privides access to:
* National Electricity Market (NEM) data
* Wholesale Electricity Market (WEM) data
* Australian PV Institute (APVI) data

{opennemr} provides a simple interface for R to access the public API provided
to OpenNEM. Each of the endpoints documented in the API documentation 
(https://api.opennem.org.au/docs) have been implemented in opennemr. 


## Installation

{opennemr} is not yet available on CRAN.

You can install the development version of opennemr from [GitHub](https://github.com/) with:

``` r
# if you don't have devtools installed, first run:
# install.packages("devtools")
devtools::install_github("RPBatchelor/opennemr")
```

## Using {opennemr}

The public OpenNEM API (current version 3.13.2-alpha.10) provides access to 
two types of data:
- **Static lists** such as:
* list of stations and facilities
* list of networks and network regions
* list of fueltechs
* time periods and intervals
- **Statistics** such as:
* Facility power and energy data
* Interconnector flows

### Static lists

A key function of opennemr is to download static lists from {opennemr}.
This is a useful first step to build the necessary inputs for later downloading
time series statistc data.

First we'll load the {opennemr} package and the {tidyverse}

```{r network list example}
library(opennemr)
library(tidyverse)
```

Now we'll create some dataframes of the relevant static list data from the OpenNEM
database

```{r}
networks <- get_network_list()
stations <- get_station_list()
fueltechs <- get_fueltech_list()
```

### Time series statistics

Now we'll look to download some actual power data for a given station. 
From the *stations* list we know the station code for Loy Yang A is *LOYYANGA* 
and this station is on the NEM network. We can get data at 30 minute intervals 
for a period of 7 days using the below:

```{r}
lya_power <- get_power_by_station(network_code = "NEM",
                                  station_code = "LOYYANGA",
                                  interval = "30m", 
                                  period = "7d")

```

This is what the data looks like:

```{r}
str(lya_power)

```

It's now very simple to chart the power output of each generator at Loy Yan A
over time

```{r}
(p <- lya_power |> 
    ggplot(aes(x = period_start, y = value)) +
    geom_line() +
    facet_wrap(~code) + 
    theme_grey() +
    labs(x = "Period start",
         y = "Power (MW)")
)
```

The period of time over which you can download data for a given station can be 
found by running {get_period_list()}. Similarly, the intervals between data points
can be found by running {get_interval_list()}. 

**Note** The OpenNEM API developer documentation indicates that start and end 
dates can be called within the REST API, but this has not yet been build into 
the functionality of {opennem}. For now, specifying the *period* in the function
call will download data from the current date backwards by the specified duration.


We can also download all power generation data by fueltech for a given network.
Let's get all generation data for Victoria (region code *VIC1*) for the month 
of March 2023.


```{r}

vic_power_fueltech <- get_power_by_fueltech_region(network_code = "NEM",
                                                   network_region_code = "VIC1",
                                                   month = "2023-03-01")

```

**Note** The way the API works for this particular endpoint is that both
power and price data are returned

```{r}
unique(vic_power_fueltech$type)
```

Therefore,, filter for either of these variables once you've got the data. 
Below charts power by fueltech for a 1 week period

```{r}
(p <- vic_power_fueltech |> 
    filter(period_start >= "2023-03-24",
           period_start < "2023-03-31",
           type == "power") |> 
    ggplot() +
    geom_area(aes(x = period_start,
                  y = value, 
                  fill = fueltech),
              show.legend = TRUE) +
    theme_grey() +
    labs(x = "Period start",
         y = "Power (MW)",
         fill = "Fuel technology")
)
```


## Disclaimer
The {opennemr} pacakge is not associated with the OpenNEM team, and has been 
developed independently to facilitate a straightforward access to the API backend
in R. All data is provided subject to any restrictions and licensing arrangements
noted on the OpenNEM website. 










