---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# opennemr

<!-- badges: start -->
<!-- badges: end -->

The goal of opennemr is to provide simple access to the OpenNEM API via R. 
Each of the endpoints documented at https://api.opennem.org.au/docs have been
implemented in opennemr. 

## Installation

You can install the development version of opennemr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("RPBatchelor/opennemr")
```

## Example

This is a basic example which shows you how to solve a common problem:

Query the OpenNEM API to get a list of fueltechs available.

```{r network list example}
library(opennemr)
## basic example code
get_network_list()
```


Download power and price data from OpenNEM API for a given month and region

```{r power data example}
library(ggplot2)
library(tidyverse)

df <- get_power_by_fueltech_region(
  network_code = "NEM",
  network_region_code = "VIC1",
  month = "2023-03-01"
)

p <- df |> 
  filter(period_start >= "2023-03-01" & 
           period_start < "2023-03-02",
         type == "power") |> 
  ggplot() +
  geom_area(aes(x = period_start, y = data, fill = fueltech))
p

```
